The value of being indispensable: a cooperative approach to bargaining with a continuum of players

### 2024-02-08 src/paper/theory.tex
More elegant proof for proposition 7

\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{assumption}{Assumption}

@@

\begin{proposition}
    \label{prop:many_sided_weighted}
    Let $f$ be continuous on $[0, 1]^L$.
    Then,
    \begin{align*}
        \varphi_P^\infty & = \int_0^1 \lambda_P t^{\lambda_P - 1} f(t^{\lambda_1}, \dots, t^{\lambda^L}) \dt
    \end{align*}
    and
    \begin{align*}
        \varphi_{A^l}^\infty & = \int_0^1 t^{\lambda_P} \lambda_l t^{\lambda_l - 1} \partial_l f(t^{\lambda_1}, \dots, t^{\lambda^L}) \dt.
    \end{align*}
\end{proposition}

@@

\begin{lemma}
    \label{lem:convergence_to_manifold}
    Consider the random vector $X_n$ with values from $[0, 1]^L$.
    Let $S_n = \sum_l[X_n]_l / L$, with the cumulative distribution function $G_n(s)$.
    Assume that for every $s \in [0, 1]$, $X_n \mid n S_n = \lfloor ns \rfloor \xrightarrow[]{L^1} h(s)$ where $h(s)$ is some continuous function.
    Let $S_n \xrightarrow[]{d} S$.
    Then $X_n \xrightarrow[]{d} h(S)$.
\end{lemma}

\begin{proof}[Proof of \cref{lem:convergence_to_manifold}]
    We have to show that
    \begin{align*}
        \lim_{n \to \infty} \Pr(X_n \leq x) = \Pr(h(S) \leq x).
    \end{align*}
    for any $x$ where the cdf of $h(S)$ is continuous.

    Let us start by conditioning on $S_n$:
    \begin{align*}
        \Pr(X_n \leq x) = \int_0^1 \Pr(X_n \leq x \mid S_n = s) \dG_n(s).
    \end{align*}
    Now consider the following:
    \begin{align*}
        \lim_{n \to \infty} \left| \Pr(X_n \leq x) - \Pr(h(S) \leq x) \right| &= \lim_{n \to \infty} \left| \int_0^1 \Pr(X_n \leq x \mid S_n = s) \dG_n(s) - \Pr(h(S) \leq x) \right| \\
        &\leq \lim_{n \to \infty} \left| \int_0^1 \Pr(X_n \leq x \mid S_n = s) \dG_n(s) - \Pr(h(S_n) \leq x) \right| \\
        &+ \lim_{n \to \infty} \left| \Pr(h(S_n) \leq x) - \Pr(h(S) \leq x) \right|
    \end{align*}
    for all $x \in [0, 1]^L$.
    I will show that both limits on the right-hand side are zero.

    Let us start with the first term.
    \begin{align*}
        \lim_{n \to \infty} &\left| \int_0^1 \Pr(X_n \leq x \mid S_n = s) \dG_n(s) - \Pr(h(S_n) \leq x) \right| \\
        &= \lim_{n \to \infty} \left| \int_0^1 \Pr(X_n \leq x \mid S_n = s) \dG_n(s) - \int_0^1 \mathbf{1}[h(s) \leq s] \dG_n(s) \right| \\
        &= \lim_{n \to \infty} \left| \int_0^1 \Pr(X_n \leq x \mid S_n = s) - \mathbf{1}[h(s) \leq x] \dG_n(s) \right| \\
        &= \lim_{n \to \infty} \left| \int_0^1 \underbrace{\Pr(X_n \leq x \mid nS_n = \lfloor ns \rfloor) - \mathbf{1}[h(s) \leq x]}_{\coloneqq r(s)} \dG_n(s) \right| \\
        &\leq \lim_{n \to \infty} \int_0^1 \left| \Pr(X_n \leq x \mid nS_n = \lfloor ns \rfloor) - \mathbf{1}[h(s) \leq x] \right| \dG(s) \\
        &+ \lim_{n \to \infty} \left| \int_0^1 r(s) \dG_n(s) - \int_0^1 r(s) \dG(s) \right| \\
        &= 0.
    \end{align*}
    The first limit is zero because $X_n \mid n S_n = \lfloor ns \rfloor \xrightarrow[]{L^1} h(s)$.
    The second limit is zero because of the continuous mapping theorem, as $S_n \xrightarrow[]{p} S$ and $\Pr_S(s : r\text{ is not continuous at }s)= 0$.

    All that remains is to show that the second term is also zero.
    This simply follows from the continuous mapping theorem, as $f$ is continuous and $S_n \xrightarrow[]{d} S$.
\end{proof}

@@

    Let us first define the random variable $S_n = \sum_l[X_n]_l / L$, the total number of small players placed before $P$.
    Also, define $h(s) = (s, \dots, s)$.
    Now let us look at the conditional distribution $n X_n | n L S_n = k$.
    As all permutations have equal probability, this is the same as picking $kL$ firms from the set $N \setminus \{P\}$ without replacement.
    The numbers of each type in this sample will follow a multinomial distribution with $kL$ trials, $L$ events and probabilities $1/L$.

    Let us now show that $\Pr(X_n < x \mid S_n = s) \xrightarrow[]{\mathrm{u}} \mathbf{1}[(s, \dots, s) \leq x]$ on $s \in \mathrm{supp}(S_n)$.

    First, observe that
    \begin{align*}
        [\E(n X_n | nS_n = \lfloor ns \rfloor)]_l &= \lfloor ns \rfloor, \\
        [\Var(n X_n | nS_n = \lfloor ns \rfloor)]_l &= \lfloor ns \rfloor \frac{L-1}{L}
    \end{align*}
    for all $\quad l = 1, \dots, L$.
    Consequently,
    \begin{align*}
        [\E(X_n | nS_n = \lfloor ns \rfloor)]_l &= \frac{\lfloor ns \rfloor}{n}, \\
        [\Var(X_n | nS_n = \lfloor ns \rfloor)]_l &= \frac{\lfloor ns \rfloor}{n^2} \frac{L-1}{L}
    \end{align*}
    for all $\quad l = 1, \dots, L$.

    Notice that, as $n \to \infty$, $[\Var(X_n | nS_n = \lfloor ns \rfloor)]_l \to 0$.
    Furthermore, $\E(X_n | nS_n = \lfloor ns \rfloor) \to h(s)$.
    Thus, for any $s \in [0, 1], X_n | nS_n = \lfloor ns \rfloor \xrightarrow[]{L^2} h(s)$.

    Let us now turn to the random variable $S_n$.
    Observe that because the probability of each permutation is the same, player $P$ has the same probability of being in any position.
    In other words, $S_n$ follows a uniform distribution on $0, \frac{1}{nL}, \frac{2}{nL} \dots, 1$.
    Consequently, $S_n$ converges in distribution to the uniform distribution $U[0, 1]$.

    I have now shown that the assumptions of \cref{lem:convergence_to_manifold} are satisfied.
    % TODO

    Then, by \cref{lem:many_sided_manifold}, the Shapley-value is
    \begin{align*}
        \varphi_P^\infty & = \int_0^1 f(t, \dots, t) \dt.
    \end{align*}

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{assumption}{Assumption}

@@

\begin{theorem}
    \label{prop:many_sided_weighted}
    Let $f$ be continuous on $[0, 1]^L$.
    Then,
    \begin{align*}
        \varphi_P^\infty & = \int_0^1 \lambda_P t^{\lambda_P - 1} f(t^{\lambda_1}, \dots, t^{\lambda^L}) \dt
    \end{align*}
    and
    \begin{align*}
        \varphi_{A^l}^\infty & = \int_0^1 t^{\lambda_P} \lambda_l t^{\lambda_l - 1} \partial_l f(t^{\lambda_1}, \dots, t^{\lambda^L}) \dt.
    \end{align*}
\end{theorem}

@@

\begin{lemma}
    \label{lem:convergence_to_manifold}
    Consider the random vector $X_n$ with values from $[0, 1]^L$.
    Let $S$, be a random variable with support $[0, 1]$ and cumulative distribution function $G_n(s)$.
    Assume that for every $s \in [0, 1]$, $X_n \mid S = s \xrightarrow[]{a.s.} h(s)$ where $h(s)$ is some continuous function.
    Then $X_n \xrightarrow[]{d} h(S)$.
\end{lemma}

\begin{proof}[Proof of \cref{lem:convergence_to_manifold}]
    We have to show that
    \begin{align*}
        \lim_{n \to \infty} \Pr(X_n \leq x) = \Pr(h(S) \leq x).
    \end{align*}

    First, note that

    Let us start by conditioning on $S$:
    \begin{align*}
        \Pr(X_n \leq x) = \int_0^1 \Pr(X_n \leq x \mid S = s) \dG(s).
    \end{align*}
    Now consider the following:
    \begin{align*}
        \lim_{n \to \infty} \left| \Pr(X_n \leq x) - \Pr(h(S) \leq x) \right| &= \lim_{n \to \infty} \left| \int_0^1 \Pr(X_n \leq x \mid S = s) \dG(s) - \Pr(h(S) \leq x)\right| \\
        &= \lim_{n \to \infty} \left| \int_0^1 \Pr(X_n \leq x \mid S = s) \dG(s) - \int_0^1 \mathbf{1}[h(s) \leq 1] \dG(s) \right| \\
        &= \lim_{n \to \infty} \left| \int_0^1 \Pr(X_n \leq x \mid S = s) - \mathbf{1}[h(s) \leq x] \dG(s) \right| \\
        &\leq \lim_{n \to \infty} \int_0^1 \left| \Pr(X_n \leq x \mid S = s) - \mathbf{1}[h(s) \leq x] \right| \dG(s) \\
        &= \int_0^1 \lim_{n \to \infty} \left| \Pr(X_n \leq x \mid S = s) - \mathbf{1}[h(s) \leq x] \right| \dG(s) \\
        &= 0
    \end{align*}
    for all $x \in [0, 1]^L$.
    The exchange of the limit and the integral is justified by the fact that the integrand is dominated by a constant function, and therefore the dominated convergence theorem applies.
    The last equality follows from the fact that $X_n \mid S = s \xrightarrow[]{a.s.} h(s)$. 
\end{proof}

@@

    First, let us switch to an alternative formulation of the problem.
    Remember, that the probability of each permutation is the same.
    Therefore, the following procedure leads to the same distribution over permutations: (1) for each player, draw a random number from the uniform distribution on $[0, 1]$, then sort the players according to the drawn numbers.
    This is equivalent to the original problem, as the probability of each permutation is the same.

    Now let us characterize $X_n$ in terms of this procedure.
    $[X_n]_l$ is the proportion of players of type $A_l$ that are placed before player $P$, or, equivalently, the number of players of type $A_l$ that have drawn a number less than player $P$.
    I.e.,
    \begin{align*}
        [X_n]_l = \frac{1}{n}\sum_{i = 1}^n \mathbf{1}[U_{li} < U_P].
    \end{align*}
    For any finite $n$, we can disregard ties, as they have probability zero.

    Now let us condition this probability distriibution on the draw of player $P$:
    \begin{align*}
        [X_n]_l < x \mid U_P = u = \frac{1}{n} \sum_{i=1}^n \mathbf{1}[U_{li} < u].
    \end{align*}
    As the draws are independent, by the strong law of large numbers,
    \begin{align*}
        [X_n]_l < x \mid U_P \xrightarrow[a.s.]{} \E[\mathbf{1}[U_{li} < u]] = u.
    \end{align*}

    Now we can use \cref{lem:convergence_to_manifold} to conclude that $X_n \xrightarrow[]{d} (U, \dots, U)$.
    Finally, by \cref{lem:many_sided_manifold}, the Shapley-value is
    \begin{align*}
        \varphi_P^\infty & = \int_0^1 f(t, \dots, t) \dt.
    \end{align*}
